---
title: "Colombian geographic reference database:  database creation"
author: "Marius Bottin"
format: 
  gfm:
    df-print: kable
bibliography: sources.bib
---

The files have been downloaded with the code from <./step1_downloading_sources.md>.
Now it is time to create the database and to start analysing the maps.

This might change in the future, but I believe that we will first load the shapefiles in R, then write them in a simple spatialite database.

The advantage of spatialite database is that QGIS manages them natively without any need for additional software, R manages them pretty well too.
Additionally, a spatialite database is a simple file, that may be shared quite easily, or even integrated in a R-package or a python module.

# Reading files

```{r}
files_sourceMaps<-dir("./sourceMaps",recursive = T)
bs_files_sourceMaps<-basename(files_sourceMaps)
dir_files_sourceMaps<-dirname(files_sourceMaps)
downloaded<-read.csv("./sourceMaps/sourcesDownloaded.csv",row.names=1)
stopifnot(downloaded$testFile%in%bs_files_sourceMaps)
mainFiles<-paste0("./sourceMaps/",files_sourceMaps[match(downloaded$testFile,bs_files_sourceMaps)])
```


Now we can read the files and put them in a list `data`.
We will transform the data and put them into MAGNAS-SIRGAS, more fitted to data in Colombia.


```{r}
require(sf)
data<-lapply(mainFiles,st_read)
names(data)<-downloaded$name
magnas<-st_crs(4996)
data<-lapply(data,st_transform,crs=magnas)
```


# Creating the database and putting the raw data in it

With the `st_write` function of the `sf` package, creating a spatialite database is relatively simple.
Simply putting a new layer in the spatialite file would create the database if it does not exist yet.


```{r}
require(RSQLite)
dbDir<-"./database/"
if(!dir.exists(dbDir)){dir.create(dbDir)}
spatialite_file<-paste0(dbDir,"/geogref.sqlite3")
nameTables<-tolower(paste0("raw_", downloaded$name))
for(i in 1:length(data))
{
  st_write(data[[i]],dsn=spatialite_file,layer=nameTables[i],driver="SQLite",append=F)
}
```


¡So now we have the database with the raw data in it!
Let's play a bit with it

# Analysing raw spatial tables

In order to be able to play with the raw tables in the spatialite database, it may be useful to create a connection to it in the R session:

```{r}
geogref<-dbConnect(SQLite(),spatialite_file,loadable.extensions=TRUE)
dbGetInfo(geogref)[["loadableExtensions"]]
```

In order to make operations go faster, we need spatial indexing on the 3 tables

```{r}
dbListTables(geogref)
geoms<-dbReadTable(geogref,"geometry_columns")
m_tabname<-match(geoms$f_table_name,nameTables)
indexesToCreate<-paste0(geoms$f_table_name,"_geom_idx")
(indexes<-dbGetQuery(geogref,"SELECT * FROM SQLIte_master WHERE type='index'"))
already<-indexesToCreate%in%indexes$name
if(sum(!already))
{
  for(i in (1:length(indexesToCreate))[!already])
  {
    dbExecute(geogref,paste0("CreateSpatialIndex ('", geoms$f_table_name[i], "', '",geoms$f_geometry_column[i],"')"))
  }
}
```

```{r}
dbGetQuery(geogref,"SELECT rv1.ogc_fid,rv2.ogc_fid FROM raw_veredas rv1 JOIN  raw_veredas rv2 ON ST_Intersects(rv1.GEOMETRY,rv2.GEOMETRY)=1 AND rv1.ogc_fid != rv2.ogc_fid")
```


Plot the unified ("dissolved") polygons from the 3 tables

```{r}

```





# Turning off the light and leaving

```{r}
dbDisconnect(geogref)
```

